---
title: "Practice Midterm 1"
author: "Gregory Schlett"
date: "03-04-2025"
output: html_document
---

#Practice Midterm 1 - Naive or ETS?

```{r}
getwd()
```


```{r}
library(fpp2)

data("arrivals")

#Here I pull US Arrivals
us_arrivals <- arrivals[, "US"]

#This plots the time series
autoplot(us_arrivals) + ggtitle("US Arrivals to Australia") + ylab("Arrivals") + xlab("Year")
```
###### Arrivals is increasing over time 

```{r}
#Summary and Boxplot
summary(us_arrivals)
boxplot(us_arrivals, main="Boxplot of US Arrivals to Australia", ylab="Arrivals")
```
```{r}
#Decomposition
us_arrivals_decomp <- decompose(us_arrivals, type="multiplicative")
autoplot(us_arrivals_decomp)
```
###### Strong Seasonality - repeating patterns
###### Multiplicative - seasonal variations are proportional to the trend
###### Summer months have around the peaks

```{r}
#Seasonal Adjustment
us_arrivals_adj <- seasadj(us_arrivals_decomp)
autoplot(us_arrivals_adj) +
  autolayer(us_arrivals, series="Actual", alpha=0.5) +
  ggtitle("Seasonally Adjusted US Arrivals")
```

```{r}
#Naive
fit_naive <- naive(us_arrivals, h=12)

autoplot(residuals(fit_naive)) + ggtitle("Residuals of Naïve Forecast")

hist(residuals(fit_naive), main="Histogram of Residuals", xlab="Residuals")

# Fitted values vs. residuals
plot(fitted(fit_naive), residuals(fit_naive), main="Fitted Values vs. Residuals")

# Actual values vs. residuals
plot(us_arrivals, residuals(fit_naive), main="Actual Values vs. Residuals")

Acf(residuals(fit_naive))
```
###### Residuals are not perfectly normal
##### Uncorrelated residuals - successful Naive

```{r}
#Accuracy Measure
accuracy(fit_naive)
```
###### ME = close to 0 ... no strong bias
###### High RMSE - predictions less accurate
###### Low MAE
###### MAPE around 10% - reasonable accuracy
###### MASE > 1 - Model performs worse than Naive (in this case slight)
###### ACF1 close to 0 - no strong autocorrelation

```{r}
#Forecast
future_forecast <- forecast(fit_naive, h=12)
autoplot(future_forecast) + ggtitle("Forecast for Next Year")
```
###### As time progresses uncertainty becomes exponentially greater
###### Seasonality and trend are not fully captured
###### Flat Projection - future values will remain around the most recent historical value
###### Not the best model choice - Let's Try Exponential Smoothing!

```{r}
#Exponential Smoothing
library(forecast)

# Fit ETS model
fit_ets <- ets(us_arrivals)

summary(fit_ets)

checkresiduals(fit_ets)

# Forecast for next year
forecast_ets <- forecast(fit_ets, h=12)

autoplot(forecast_ets)
```
###### Shows stable trend with some seasonality 
###### Exponential Smoothing appears to work much better
