---
title: "Practice Midterm 1 & 2"
author: "Gregory Schlett"
date: "03-04-2025"
output: html_document
---

#Practice Midterm 1 & 2 

```{r}
getwd()
```


```{r}
library(fpp2)

data("arrivals")

#Here I pull US Arrivals
us_arrivals <- arrivals[, "US"]

#This plots the time series
autoplot(us_arrivals) + ggtitle("US Arrivals to Australia") + ylab("Arrivals") + xlab("Year")
```
###### Arrivals is increasing over time 

```{r}
#Summary and Boxplot
summary(us_arrivals)
boxplot(us_arrivals, main="Boxplot of US Arrivals to Australia", ylab="Arrivals")
```
```{r}
#Decomposition
us_arrivals_decomp <- decompose(us_arrivals, type="multiplicative")
autoplot(us_arrivals_decomp)
```
###### Strong Seasonality - repeating patterns
###### Multiplicative - seasonal variations are proportional to the trend
###### Summer months have around the peaks

```{r}
#Seasonal Adjustment
us_arrivals_adj <- seasadj(us_arrivals_decomp)
autoplot(us_arrivals_adj) +
  autolayer(us_arrivals, series="Actual", alpha=0.5) +
  ggtitle("Seasonally Adjusted US Arrivals")
```

```{r}
#Naive
fit_naive <- naive(us_arrivals, h=12)

autoplot(residuals(fit_naive)) + ggtitle("Residuals of Naïve Forecast")

hist(residuals(fit_naive), main="Histogram of Residuals", xlab="Residuals")

# Fitted values vs. residuals
plot(fitted(fit_naive), residuals(fit_naive), main="Fitted Values vs. Residuals")

# Actual values vs. residuals
plot(us_arrivals, residuals(fit_naive), main="Actual Values vs. Residuals")

Acf(residuals(fit_naive))
```
###### Residuals are not perfectly normal
##### Uncorrelated residuals - successful Naive

```{r}
#Accuracy Measure
accuracy(fit_naive)
```
###### ME = close to 0 ... no strong bias
###### High RMSE - predictions less accurate
###### Low MAE
###### MAPE around 10% - reasonable accuracy
###### MASE > 1 - Model performs worse than Naive (in this case slight)
###### ACF1 close to 0 - no strong autocorrelation

```{r}
#Forecast
future_forecast <- forecast(fit_naive, h=12)
autoplot(future_forecast) + ggtitle("Forecast for Next Year")
```
###### As time progresses uncertainty becomes exponentially greater
###### Seasonality and trend are not fully captured
###### Flat Projection - future values will remain around the most recent historical value
###### Not the best model choice - Let's Try Exponential Smoothing!

```{r}
#Exponential Smoothing
library(forecast)

# Fit ETS model
fit_ets <- ets(us_arrivals)

summary(fit_ets)

checkresiduals(fit_ets)

# Forecast for next year
forecast_ets <- forecast(fit_ets, h=12)

autoplot(forecast_ets)
```
###### Shows stable trend with some seasonality 
###### Exponential Smoothing appears to work much better

```{r}
library(fpp2)
library(TTR)         
library(ggplot2)     
library(forecast)    

data("arrivals")

us_arrivals <- arrivals[, "US"]

autoplot(us_arrivals) +
  labs(title = "US Arrivals to Australia",
       x = "Year", Y = "Arrivals") + 
  theme_minimal()

sma_3 <- SMA(us_arrivals, n = 3)
sma_6 <- SMA(us_arrivals, n = 6)
sma_9 <- SMA(us_arrivals, n = 9)

sma_df <- data.frame(
  Time = as.numeric(time(us_arrivals)), 
  Value = as.numeric(us_arrivals), 
  sma_3 = as.numeric(sma_3),
  sma_6 = as.numeric(sma_6),
  sma_9 = as.numeric(sma_9)
)

ggplot(sma_df, aes(x = Time)) +
  geom_line(aes(y = Value), color = "black", linewidth = 1) + 
  geom_line(aes(y = sma_3), color = "red", linewidth = 1) +   
  geom_line(aes(y = sma_6), color = "blue", linewidth = 1) +  
  geom_line(aes(y = sma_9), color = "green", linewidth = 1) + 
  labs(title = "US Arrivals to Australia with Moving Averages",
       x = "Year", y = "Arrivals") +
  theme_minimal()

best_sma <- sma_6
best_sma_ts <- ts(best_sma, start = start(us_arrivals), frequency = 12)

#Naive based on SMA
forecast_sma <- forecast(best_sma_ts, h = 12)

autoplot(forecast_sma) +
  labs(title = "12-Month Forecast of US Arrivals using SMA(6)",
       x = "Year", y = "Forecasted Arrivals") +
  theme_minimal()
```

```{r}
# Fit the Holt-Winters model
hw_model <- hw(us_arrivals, seasonal = "multiplicative", h = 12)

summary(hw_model)
```

```{r}
alpha_value <- hw_model$model$par["alpha"]
beta_value <- hw_model$model$par["beta"]
gamma_value <- hw_model$model$par["gamma"]
sigma_value <- sqrt(hw_model$model$sigma2)

initial_values <- hw_model$model$initstate

cat("Alpha:", alpha_value, "\n")
cat("Beta:", beta_value, "\n")
cat("Gamma:", gamma_value, "\n")
cat("Sigma:", sigma_value, "\n")
cat("Initial Level:", initial_values[1], "\n")
cat("Initial Trend:", initial_values[2], "\n")
cat("Initial Seasonality:", initial_values[-(1:2)], "\n")

residuals_hw <- residuals(hw_model)

autoplot(residuals_hw) +
  ggtitle("Residuals of Holt-Winters Model") +
  theme_minimal()

hist(residuals_hw, breaks = 20, main = "Histogram of Residuals", col = "blue")

autoplot(fitted(hw_model), residuals_hw) +
  ggtitle("Fitted Values vs. Residuals")

autoplot(us_arrivals, residuals_hw) +
  ggtitle("Actual Values vs. Residuals")

Acf(residuals_hw, main = "ACF of Residuals")
```

